### 2020-07-20 abstract.tex
final comments and edits from katy

Molten salt reactors is a class of advanced nuclear reactors that promise
numerous improvements over the current fleet of largely light-water reactors.
As the world continues its transition towards low-carbon electricity
generation to combat climate change, molten salt reactors is a potential
option in the near-term future for replacing fossil fuel and ageing nuclear
power plants. At the current state of development, molten salt reactors still
require extensive research to become viable. This thesis presents the latest
developments in Moltres, a simulation tool for molten salt reactors. These
new developments are: support for coupling the incompressible Navier-Stokes
and the delayed neutron precursor looping systems, and a model for
simulating decay heat from fission products at steady state and during
transients. This work demonstrates these capabilities through multiphysics
simulations of the Molten Salt Fast
Reactor concept. This work first verifies the six-group neutron diffusion
results from Moltres against continuous-energy Monte Carlo neutron transport
results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
the six-group neutron energy spectra from Moltres agreed with the high
fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
small discrepancies on the order of 100 pcm, which is smaller than the
$-256.7$ pcm discrepancy reported in literature with the same six-group
neutron diffusion approach. The decay heat model shows an expected flattening
of the temperature distribution due to the movement and dispersion of the
decay heat precursors throughout the primary coolant loop. This work also
demonstrates and verifies steady state and
transient multiphysics simulations of the Molten Salt Fast Reactor. The
transient scenarios under study were unprotected instances of reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
state and transient results were verified against results from another paper
that presented results for the same cases. The steady-state temperature and
velocity distributions, and the peak neutron flux showed good agreement with
the literature results. Minor differences in the delayed neutron precursor
distribution and the in-core delayed neutron fraction were explainable with
the differences in the handling of turbulence in the models. In three of the
transient results (reactivity insertion, loss of heat sink, and pump
overspeed), Moltres reproduced the expected magnitude and pattern of the
reactor response to these transient initiators.
The loss of flow results showed greater discrepancies that are attributed
to differences in the fluid dynamics modeling in Moltres and the other models.
Through the verification studies, this work has also identified avenues
for further Moltres software development.

>>>>

Molten salt reactors, a class of advanced nuclear reactors, promise
numerous improvements over the current fleet of largely light-water reactors.
As the world continues its transition towards low-carbon electricity
generation to combat climate change, molten salt reactors is a potential
option in the near-term future for replacing fossil fuel and ageing nuclear
power plants. At the current state of development, molten salt reactors still
require extensive research to become viable. This thesis presents the latest
developments in Moltres, a simulation tool for molten salt reactors. These
new developments are: support for coupling the incompressible Navier-Stokes
and the delayed neutron precursor looping systems, and a model for
simulating decay heat from fission products at steady state and during
transients. This work demonstrates these capabilities through multiphysics
simulations of the Molten Salt Fast
Reactor concept. This work first verifies the six-group neutron diffusion
results from Moltres against continuous-energy Monte Carlo neutron transport
results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
the six-group neutron energy spectra from Moltres agreed with the high
fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
small discrepancies on the order of 100 pcm, which is smaller than the
$-256.7$ pcm discrepancy reported in literature with the same six-group
neutron diffusion approach. The decay heat model shows an expected flattening
of the temperature distribution due to the movement and dispersion of the
decay heat precursors throughout the primary coolant loop. This work also
demonstrates and verifies steady state and
transient multiphysics simulations of the Molten Salt Fast Reactor. The
transient scenarios under study include unprotected instances of reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed. This thesis
verified the steady state and transient results against data from the
literature for the same case studies. The steady-state temperature and
velocity distributions, and the peak neutron flux showed good agreement with
the literature results. Minor differences in the delayed neutron precursor
distribution and the in-core delayed neutron fraction were explainable with
the differences in the handling of turbulence in the models. In three of the
transient results (reactivity insertion, loss of heat sink, and pump
overspeed), Moltres reproduced the expected magnitude and pattern of the
reactor response to these transient initiators.
The loss of flow results showed greater discrepancies that are attributed
to differences in the fluid dynamics modeling in Moltres and the other models.
Through the verification studies, this work has also identified avenues
for further Moltres software development.

### 2020-07-20 model.tex
final comments and edits from katy

This work demonstrates the \gls{MSFR} simulation capabilities of Moltres, a
multiphysics simulation tool for \glspl{MSR} \cite{lindsay_introduction_2018}.
In particular, this work introduces two new capabilities: full support
for coupling incompressible flow with the existing delayed neutron precursor
looping capability, and a decay heat model. The former allows users to
simulate non-trivial flow patterns in the core and simultaneously loop the
precursors through an external region, and the latter to simulate delayed
heating from fission products.
To run simulations with Moltres, the user must provide input
group constant data from a neutron transport solver for the
multigroup neutron diffusion calculations, and a mesh file representing the
geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
for the former and Trelis/Cubit \cite{noauthor_trelis_2018} for the latter.
This chapter provides brief introductions to
Serpent 2, \gls{MOOSE}, and Moltres, and the modeling approach for the
\gls{MSFR} multiphysics simulations in this thesis.

@@

\gls{MOOSE} itself relies on libMesh \cite{kirk_libmesh:_2006} and
PETSc \cite{satish_petsc_2019} for mesh handling and \gls{PDE} solver
functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
and automatic variable scaling amongst other advanced featuresin finite
element analysis. Full
coupling is maintained by the execution of Newton-based solves on the
weak formulations of the multiple \glspl{PDE} to minimize the residual values.
Fully-coupled solves are essential for accurately resolving systems with
strongly interacting physics. The \gls{MSR} concept is one such example, in
which the neutronics and thermal-hydraulics are tightly coupled through the
Doppler effect and the temperature dependence of liquid fuel salt density.

@@

Table \ref{table:corebc} summarizes the boundary conditions for all variables
on all of the relevant boundaries. Figure \ref{fig:msfrbc} shows the locations
of the various boundaries listed in the table. This thesis introduces the
CoupledOutflow boundary condition for $C_i$ and $\omega_j$ that allows users
to couple these variables to the outlet velocity components (e.g. $u_x$,
$u_y$). Without this new feature, users could only use uniform or fixed
function-based velocity profiles in conjunction with the precursor looping
capability.

@@

Moltres also accounts for the decay of
\glspl{DNP} outside the central core region by simulating its flow in a
separate 1-D pipe geometry. This outer loop pipe calculation is implicitly
coupled to the active core simulation through Picard iterations in MOOSE's
MultiApp functionality and inlet/outlet boundary values. For this work with
the \gls{MSFR} model, the pipe length is 2.255 m with salt flowing at 1.1275 m
s$^{-1}$ for an out-of-core residence time of 2 s. The present author derived
these parameters from the reference specifications of 4 s cycle
time and 50\% out-of-core salt fraction (Table \ref{table:msfr}).

@@

A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region
to maintain the nominal 2 s out-of-core residence time. The governing equation
for temperature, derived from equation \ref{eq:temp}, is:
%
\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p u
    \frac{\partial T}{\partial x} - Q_{hx} \label{eq:temploop} \\
    Q_{hx} &= \alpha (T - T_i) \delta (x_0) \label{eq:hx} \\
    \intertext{where}
    Q_{hx} &= \text{heat removal rate through the heat exchanger [W],} 
    \nonumber \\
    \alpha &= \text{heat transfer coefficient [W$\cdot$K$^{-1}$],} \nonumber
    \\
    T_i &= \text{temperature of the intermediate salt [K],} \nonumber \\
    x_0 &= \text{position of the point heat exchanger [m].} \nonumber
\end{align}

@@

Table \ref{table:loopbc} summarizes the boundary conditions for all variables
on the inlet and outlet of the 1-D outer loop region. The inlet boundary
conditions are all Dirichlet boundary conditions. The inlet boundary values
are set by the outflow from the central core region that this inlet is
connected to in the actual reactor geometry. The outlet boundary conditions
are all outflow boundary conditions as shown in the table.

>>>>

This work demonstrates the \gls{MSFR} simulation capabilities of Moltres, a
multiphysics simulation tool for \glspl{MSR} \cite{lindsay_introduction_2018}.
In particular, this work introduces two new capabilities: full support
for coupling incompressible flow with the existing delayed neutron precursor
looping capability, and a decay heat model. The former allows users to
simulate non-trivial flow patterns in the core and simultaneously loop the
precursors through an external region, and the latter to simulate delayed
heating from fission products.
To run simulations with Moltres, the user must provide input
group constant data from a neutron transport solver for the
multigroup neutron diffusion calculations, and a mesh file representing the
geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
for the former and Trelis/CUBIT \cite{noauthor_trelis_2018} for the latter.
This chapter provides brief introductions to
Serpent 2, \gls{MOOSE}, and Moltres, and the modeling approach for the
\gls{MSFR} multiphysics simulations in this thesis.

@@

\gls{MOOSE} itself relies on libMesh \cite{kirk_libmesh:_2006} and
PETSc \cite{satish_petsc_2019} for mesh handling and \gls{PDE} solver
functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
and automatic variable scaling amongst other advanced features in finite
element analysis. Full
coupling is maintained by the execution of Newton-based solves on the
weak formulations of the multiple \glspl{PDE} to minimize the residual values.
Fully-coupled solves are essential for accurately resolving systems with
strongly interacting physics. The \gls{MSR} concept is one such example, in
which the neutronics and thermal-hydraulics are tightly coupled through the
Doppler effect and the temperature dependence of liquid fuel salt density.

@@

Table \ref{table:corebc} summarizes the boundary conditions for all variables
on all of the relevant boundaries. Figure \ref{fig:msfrbc} shows the locations
of the various boundaries listed in the table. The CoupledOutflow boundary
condition for $C_i$ and $\omega_j$ is a new feature in Moltres that allows
users to couple these variables to the outlet velocity components (e.g. $u_x$,
$u_y$). Without this boundary condition, users could only use uniform or fixed
function-based velocity profiles in conjunction with the precursor looping
capability.

@@

Moltres also accounts for the decay of
\glspl{DNP} outside the central core region by simulating its flow in a
separate 1-D pipe geometry. This outer loop pipe calculation is implicitly
coupled to the active core simulation through Picard iterations in MOOSE's
MultiApp functionality and inlet/outlet boundary values. For this work with
the \gls{MSFR} model, the pipe length is 2.255 m with salt flowing at 1.1275 m
s$^{-1}$ for an out-of-core residence time of 2 s. The present author derived
these parameters from the reference specifications of 4s cycle
time and 50\% out-of-core salt fraction (Table \ref{table:msfr}).

@@

A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region
to maintain the nominal 2s out-of-core residence time. The governing equation
for temperature, derived from equation \ref{eq:temp}, is:
%
\begin{align}
    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p u
    \frac{\partial T}{\partial x} - Q_{hx} \label{eq:temploop} \\
    Q_{hx} &= \alpha (T - T_i) \delta (x_0) \label{eq:hx} \\
    \intertext{where}
    Q_{hx} &= \text{heat removal rate through the heat exchanger [W],} 
    \nonumber \\
    \alpha &= \text{heat transfer coefficient [W$\cdot$K$^{-1}$],} \nonumber
    \\
    T_i &= \text{temperature of the intermediate salt [K],} \nonumber \\
    x_0 &= \text{position of the point heat exchanger [m].} \nonumber
\end{align}

@@

Table \ref{table:loopbc} summarizes the boundary conditions for all variables
on the inlet and outlet of the 1-D outer loop region. The inlet boundary
conditions are all Dirichlet boundary conditions. The inlet boundary values
are set by the outflow from the central core region that this inlet is
connected to in the actual reactor geometry. The outlet boundary conditions
are all outflow boundary conditions as shown in Table \ref{table:loopbc}.

### 2020-07-20 msr.tex
final comments and edits from katy

Two other \gls{MSR} designs emerged in the 1970s: the \gls{DMSR} from
\gls{ORNL} \cite{gehin_liquid_2016} and the \gls{MCFR} from the UK Atomic
Energy Authority \cite{smith_assessment_1974}. \gls{ORNL}
researchers developed the \gls{DMSR}, named after its planned use of denatured
$^{235}$U fuel (i.e. low-enriched uranium), as a once-through converter system
without fuel reprocessing in response to fuel reprocessing restrictions
introduced by President Ford in 1976. On the other hand, UK researchers
developed the lead-cooled \gls{MCFR} to enhance $^{239}$Pu breeding with the
relatively fast neutron spectrum. Both designs failed to gain traction and did
not inspire any working prototypes.

Following a lull lasting through the late 20th century, researchers at
\gls{CNRS} began research into \glspl{MSR} in 1997
\cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
performed parametric studies on the design based on safety, breeding, and
other performance metrics \cite{mathieu_thorium_2006}. Their efforts
culminated in the \gls{MSFR} concept, a fast-spectrum breeder \gls{MSR}
designed to run on the thorium fuel cycle \cite{merle_optimized_2007}. In
2008, the Generation IV International Forum highlighted the \gls{MSFR} among
other \gls{MSR} designs for further development \cite{gif_generation_2008}.
The \gls{MSFR} has also benefited from collaborative research through two
European-funded projects, the EVOL and SAMOFAR projects
\cite{euratom_final_2015}. Today, numerous other national and commercial
bodies are also developing their own \gls{MSR} designs.

@@

The various \gls{MSR} designs under development today illustrate the
flexibility of this reactor concept. Graphite-moderated thermal-spectrum
\glspl{MSR} are typically straightforward \gls{LEU} burners or
$^{232}$Th/$^{233}$U iso-breeders/breeders, while epithermal- and
fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs. Two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture. Examples of one-fluid designs include the Integral Molten
Salt Reactor from Terrestrial Energy \cite{leblanc_integral_2015} and the
\gls{MSR} design from Transatomic Power
\cite{transatomic_power_corporation_technical_2016} while two-fluid designs
include the \gls{MSFR} \cite{merle_optimized_2007} and \gls{MOSART}
\cite{ignatiev_molten_2014}. The current work involves simulations of
the \gls{MSFR} in Moltres and Section \ref{sec:msfr} describes this reactor
design in further detail.

@@

@@

@@

Both DALTON and the COMSOL-based tool featured again in a later paper by
Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D axisymmetric
model of the \gls{MSFR}. The authors, from Politecnico di Milano and
Technical University of Delft, compared \gls{MSFR} transient safety
analysis results from both tools. They referred to the COMSOL and DALTON/HEAT
\gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
universities where each application was developed. This thesis also follows
this naming convention. With
multigroup neutron diffusion and \gls{RANS} formulations on ultra fine meshes,
both models agreed on the steady-state neutron flux, temperature,
and \gls{DNP} distributions, as well as the power responses following various
accident transient initiations.
Aufiero et al. \cite{aufiero_development_2014} concurrently developed
a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. Aufiero et al. also
provided quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.

@@

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it supports a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$; AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. Using fluoride salts instead of chloride salts results in a
relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
thermal neutron spectrum than a typical sodium fast reactor. For an initial
composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of $^{233}$U for criticality under
normal operating conditions is 2.5 mol\%. However, a neutronic benchmark
study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
different neutronics software with different nuclear data provide
different $k_{eff}$ estimates even with the same isotopic compositions and
temperature distributions. The individual co-authors adjusted the
ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
uniform temperature of 973 K for their own neutronics software
\cite{brovchenko_neutronic_2019}. This thesis
performs the same exercise to adjust the inlet and outlet temperatures to
match nominal values.

>>>>

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
developed this design in response to the fuel reprocessing restrictions
introduced by President Ford in 1976. The \gls{DMSR} would operate as a
once-through
converter system without fuel reprocessing. While the fuel consists of 19.75
\% high-assay low-enriched uranium, the initial core loading includes thorium
to boost its conversion ratio throughout its lifetime. It has a continuous
online feed consisting of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous fission products, while noble metals plate out onto the walls of
the coolant loop. The older \gls{MSBR} design had a significant drawback; the
extensive neutron damage in the graphite moderator necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by having a lower power density while maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
projected that the graphite moderator would last for the entirety of the
\gls{DMSR}'s design lifetime.

A concurrent program at the UK Atomic Energy Authority led the
development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The absence of moderators and the choice
of chloride over fluoride salt resulted in a relatively hard neutron spectrum
which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
performed some experiments to study molten salt chemistry but there were no
reactor prototypes. The UK program eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, researchers at
\gls{CNRS} began research into \glspl{MSR} in 1997
\cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
performed parametric studies based on safety, breeding, and
other performance metrics \cite{mathieu_thorium_2006}. Their efforts
culminated in the \gls{MSFR} concept, a fast-spectrum breeder \gls{MSR}
designed to run on the thorium fuel cycle \cite{merle_optimized_2007}. In
2008, the Generation IV International Forum highlighted the \gls{MSFR} among
other \gls{MSR} designs for further development \cite{gif_generation_2008}.
The \gls{MSFR} has also benefited from collaborative research through two
European-funded projects, the \gls{EVOL} \cite{euratom_final_2015} and
\gls{SAMOFAR} \cite{kloosterman_20_2017} projects. Under the \gls{EVOL}
project, researchers further optimized the \gls{MSFR} design based on
neutronic and thermal-hydraulic safety analyses. They also diversified
start-up fuel options beyond the original $^{233}$U-based fuel composition to
include $^{235}$U, $^{239}$Pu, and \gls{TRU}, and established fuel
reprocessing schemes for the \gls{MSFR}. The \gls{SAMOFAR} project, which
stated approximately 2 years after the end of the \gls{EVOL} project,
supported more comprehensive safety assessments of the reactor and the
reprocessing plant, and funded a number of experiments for validation of the
\gls{MSFR}'s safety features. This thesis presents steady-state and
transient results of the \gls{MSFR} in Moltres. Therefore, Section
\ref{sec:msfr} describes the \gls{MSFR} design in further detail.

Today, numerous other national and commercial
bodies are also developing their own \gls{MSR} designs. These include the
Integral Molten Salt Reactor from Terrestrial Energy
\cite{leblanc_integral_2015}, the Transatomic Power \gls{MSR}
\cite{transatomic_power_corporation_technical_2016}, the Russian \gls{MOSART}
\cite{ignatiev_molten_2014}, and the Chinese Thorium Molten Salt Reactor
project \cite{zhang_review_2018}.

@@

The various \gls{MSR} designs under development today illustrate the
flexibility of this reactor concept. Graphite-moderated thermal-spectrum
\glspl{MSR} are typically straightforward \gls{LEU} burners or
$^{232}$Th/$^{233}$U iso-breeders/breeders, while epithermal- and
fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs. Two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture. Examples of one-fluid designs include the Integral Molten
Salt Reactor \cite{leblanc_integral_2015} and the Transatomic Power
\gls{MSR} design \cite{transatomic_power_corporation_technical_2016} while
two-fluid designs include the \gls{MSFR} \cite{merle_optimized_2007} and
\gls{MOSART} \cite{ignatiev_molten_2014}.

@@

This subsection briefly covers other \gls{MSR} features that are not the focus
of this thesis for completeness.

@@

Economic analyses of \glspl{MSR} are preliminary at the current
stage of \gls{MSR} development. Qualitatively, many technical factors favor
\glspl{MSR} over \glspl{LWR}. Some of these factors include: smaller reactor
core size due to low operating pressures, higher thermal efficiency and
cheaper air-cooling systems due to high operating temperatures, reduction in
fuel fabrication costs, and shorter reactor shutdown periods due to online
refueling \cite{dolan_1_2017}.

@@

Both DALTON and the COMSOL-based tool featured again in a 2014 paper by
Fiorina et al. \cite{fiorina_modelling_2014} for a 2-D axisymmetric
model of the \gls{MSFR}. The authors, from Politecnico di Milano and
Technical University of Delft, compared \gls{MSFR} transient safety
analysis results from both tools. They referred to the COMSOL and DALTON/HEAT
\gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
universities where each application was developed. This thesis also follows
this naming convention. With
multigroup neutron diffusion and \gls{RANS} formulations on ultra fine meshes,
both models agreed on the steady-state neutron flux, temperature,
and \gls{DNP} distributions, as well as the power responses following various
accident transient initiations.
Aufiero et al. \cite{aufiero_development_2014} concurrently developed
a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. Aufiero et al. also
provided quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.

@@

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it supports a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$; AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. Using fluoride salts instead of chloride salts results in a
relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
thermal neutron spectrum than a typical fast-spectrum reactors. For an initial
composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of $^{233}$U for criticality under
normal operating conditions is 2.5 mol\%. However, a neutronic benchmark
study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
different neutronics software with different nuclear data provide
different $k_{eff}$ estimates even with the same isotopic compositions and
temperature distributions. The individual co-authors adjusted the
ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
uniform temperature of 973 K for their own neutronics software
\cite{brovchenko_neutronic_2019}. This thesis
performs the same exercise to adjust the inlet and outlet temperatures to
match nominal values.

### 2020-07-20 neutronics.tex
final comments and edits from katy

This chapter verifies Moltres' ability to reproduce key neutronics parameters
using group constant data from Serpent 2, which is essential for accurate
neutronics calculations in the subsequent multiphysics simulations. The model
under study is a static model of the \gls{MSFR}, i.e. no salt flow, and
uniform temperature distribution to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
verification exercise builds on the previous study by Lindsay et al.
\cite{lindsay_introduction_2018} that verified Moltres' neutronics
capabilities with a two-group neutron diffusion model of the \gls{MSRE}.

\textbf{Details of the static \gls{MSFR} model in Moltres}
\begin{itemize}
    \item No salt flow (static salt)
    \begin{center}
        $v_{salt} = 0$ m$\cdot$s$^{-1}$
    \end{center}
    \item Uniform temperature of 973 K throughout the 2D core model
    \begin{center}
        $T = 973$ K
    \end{center}
    \item Six neutron energy groups
    \begin{center}
        $G = 6$
    \end{center}
    \item Eight delayed neutron precursor groups
    \begin{center}
        $I = 8$
    \end{center}
    \item Vacuum boundary conditions on the outer boundaries of the 2D core
    mesh
    \begin{center}
        $\frac{d \phi}{dx}\big|_{\text{inflow}} = 0$ m$^{-2}\cdot$s$^{-1}$
    \end{center}
\end{itemize}

@@

Moltres solves the six-group neutron diffusion equations (Equation
\ref{eq:neut}) as a
steady-state eigenvalue problem to find the $k_{\text{eff}}$ for the static
\gls{MSFR} model. Table
\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent 2 and Moltres
at 973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. Two
main factors contribute to the small discrepancies on the order of 100 pcm
between the two applications: the accuracy of the neutron diffusion
model, and the omission of the blanket tank structural material in Moltres.
The neutron diffusion model is not as accurate as the other S$_{\text{N}}$ or
SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
Regarding the omission of the blanket tank material, the present author
replaced the 2 cm-thick structural material with blanket salt. This
replacement is partly responsible for the higher $k_{\text{eff}}$ value
calculated by Moltres as fissions occur in the blanket salt. Nevertheless, the
discrepancy is smaller
than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
diffusion methods, respectively.

@@

This work compares the $\beta$ value from Moltres to the
$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks an
adjoint calculation capability.
The difference between $\beta$ and $\beta_{\text{eff}}$ is
that $\beta$ is the unweighted delayed neutron fraction
while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted by the
adjoint neutron flux. The present author calculated $\beta$ by
taking the relative difference between the $k_{\text{eff}}$ values with and
without \glspl{DNP} in Table \ref{table:keff}. The $\beta$ and
$\beta_{\text{eff}}$ values at 973 K, shown in Table \ref{table:betaeff}, are
in good agreement with a 4.43 pcm discrepancy.

>>>>

This chapter verifies Moltres' ability to reproduce key neutronics parameters
using group constant data from Serpent 2, which is essential for accurate
neutronics calculations in the subsequent multiphysics simulations. The model
under study is a static model of the \gls{MSFR}, i.e. no salt flow, and
uniform temperature distribution to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. Table
\ref{table:static} lists relevant details of the static \gls{MSFR} model
in Moltres. This
verification exercise builds on the previous study by Lindsay et al.
\cite{lindsay_introduction_2018} that verified Moltres' neutronics
capabilities with a two-group neutron diffusion model of the \gls{MSRE}.

\begin{table}[htbp!]
    \small
	\caption{Details of the static \gls{MSFR} model in Moltres.}
	\centering
	\begin{tabular}{ l c}
		\toprule
		Detail & Mathematical description \\
        \midrule
        No salt flow (static salt) & $v_{salt} = 0$ m$\cdot$s$^{-1}$ \\
        Uniform temperature of 973 K throughout the 2D core model & $T = 973$
        K \\
        Six neutron energy groups & $G = 6$ \\
        Eight delayed neutron precursor groups & $I = 8$ \\
        Vacuum boundary conditions for neutron flux &
        $\frac{d \phi}{dx}\big|_{\text{inflow}} = 0$ m$^{-2}\cdot$s$^{-1}$ \\
		\bottomrule
	\end{tabular}
	\label{table:static}
\end{table}

@@

Moltres solves the six-group neutron diffusion equations (Equation
\ref{eq:neut}) as a
steady-state eigenvalue problem to find the $k_{\text{eff}}$ for the static
\gls{MSFR} model. Table
\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent 2 and Moltres
at 973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. Two
main factors contribute to the small discrepancies on the order of 100 pcm
between the two applications: the accuracy of the neutron diffusion
model, and the omission of the blanket tank structural material in Moltres.
The neutron diffusion model is not as accurate as the other S$_{\text{N}}$ or
SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
Regarding the omission of the blanket tank material, I
replaced the 2 cm-thick structural material with blanket salt. This
replacement is partly responsible for the higher $k_{\text{eff}}$ value
calculated by Moltres as fissions occur in the blanket salt. Nevertheless, the
discrepancy is smaller
than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
diffusion methods, respectively.

@@

This work compares the $\beta$ value from Moltres to the
$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks an
adjoint calculation capability.
The difference between $\beta$ and $\beta_{\text{eff}}$ is
that $\beta$ is the unweighted delayed neutron fraction
while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted by the
adjoint neutron flux. I calculated $\beta$ by
taking the relative difference between the $k_{\text{eff}}$ values with and
without \glspl{DNP} in Table \ref{table:keff}. The $\beta$ and
$\beta_{\text{eff}}$ values at 973 K, shown in Table \ref{table:betaeff}, are
in good agreement with a 4.43 pcm discrepancy.
